---
title: "Viral Load on Fig2"
author: "Stevin Wilson"
date: "4/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(car)
library(rstatix)
library(moments)
library(ggstatsplot)
set.seed(100)
```

### Loading `data_accessibility_table.csv`

```{r}
da_tbl <- readr::read_csv('data_accessibility_table.csv',
                          show_col_types = FALSE)

dplyr::glimpse(da_tbl)
```

```{r}
da_tbl <- da_tbl %>%
  dplyr::filter(clade %in% c("21A (Delta)", 
                             "20I (Alpha, V1)", 
                             "20J (Gamma, V3)", 
                             "20G"
                             )) %>%
  dplyr::mutate(clade = factor(clade,
                levels = c( "20G", 
                            "20I (Alpha, V1)", 
                            "20J (Gamma, V3)", 
                            "21A (Delta)"
                            )))

glimpse(da_tbl)
```


```{r}
# Ct = -1.461*ln(viral_load)+33.188

## viral_load = exp(((33.188 - Ct)/(1.461)))

da_tbl_vl <- da_tbl %>%
  dplyr::mutate(
    `Viral Load` = exp(((33.188 - `Mean Ct : N gene`)/(1.461)))
    ) %>%
  dplyr::relocate(`Viral Load`, .after = `Mean Ct : N gene`)

glimpse(da_tbl_vl)
```

Writing table to a csv file

```{r}
write_csv(da_tbl_vl,
          'data_accessibility_table_viral_load.csv')
```

##### Histogram of Residuals

```{r warning = FALSE, fig.width = 7, fig.height = 5, fig.align='center'}
res_aov_n <- aov(`Viral Load` ~ clade,
  data = da_tbl_vl
)

hist(res_aov_n$residuals)
```

The histogram is severely right skewed.


##### Shapiro-Wilk

> Null Hypothesis: Data comes from a normal distribution.

> Alternate Hypothesis: Data does not come from a normal distribution.

```{r}
shapiro.test(res_aov_n$residuals)
```

*Since p-value \< 0.05, we reject the null hypothesis. The data does not follow a normal distribution.*

*We will perform Kruskal-Wallis Test to compare the Ct values for N gene between the different SARS-CoV-2 clades.*

```{r warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5, fig.align='center'}
# With our palette
(ct_N_stats_plot <- ggbetweenstats(
  data = da_tbl_vl,
  x = clade,
  y = `Viral Load`,
  type = "nonparametric",
  xlab = "Clade",
  ylab = "Viral Load (vial copies per ul (cpu))",
  var.equal = FALSE,
  plot.type = "boxviolin",
  pairwise.comparisons = FALSE,
  results.subtitle = FALSE,
  centrality.plotting = FALSE,
  bf.message = FALSE,
  p.adjust.method = "holm",
  ggtheme = ggpubr::theme_pubr()
) +
  scale_color_manual(values = c(
    "dodgerblue2",
    "#E31A1C",
    "green4",
    "#6A3D9A"
  )) +
  labs(title = "Viral Load - All Samples") +
  scale_y_log10())
```

```{r include = FALSE}
ggsave('fig2_viral_load.png',
       ct_N_stats_plot)
```

```{r warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5, fig.align='center'}

da_tbl_vl_surv <- da_tbl_vl %>%
  dplyr::filter(order_priority == "SURVEILLANCE")

# With our palette
(ct_N_stats_plot_surv <- ggbetweenstats(
  data = da_tbl_vl_surv,
  x = clade,
  y = `Viral Load`,
  type = "nonparametric",
  xlab = "Clade",
  ylab = "Viral Load (vial copies per ul (cpu))",
  var.equal = FALSE,
  plot.type = "boxviolin",
  pairwise.comparisons = FALSE,
  results.subtitle = FALSE,
  centrality.plotting = FALSE,
  bf.message = FALSE,
  p.adjust.method = "holm",
  ggtheme = ggpubr::theme_pubr()
) +
  scale_color_manual(values = c(
    "dodgerblue2",
    "#E31A1C",
    "green4",
    "#6A3D9A"
  )) +
  labs(title = "Viral Load - Only SURVEILLANCE Samples") +
  scale_y_log10())
```

```{r include = FALSE}
ggsave('fig2_viral_load_surveillance.png',
       ct_N_stats_plot_surv)
```

```{r warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5, fig.align='center'}

da_tbl_vl_sym <- da_tbl_vl %>%
  dplyr::filter(order_priority == "SYMPTOMATIC")

# With our palette
(ct_N_stats_plot_sym <- ggbetweenstats(
  data = da_tbl_vl_sym,
  x = clade,
  y = `Viral Load`,
  type = "nonparametric",
  xlab = "Clade",
  ylab = "Viral Load (vial copies per ul (cpu))",
  var.equal = FALSE,
  plot.type = "boxviolin",
  pairwise.comparisons = FALSE,
  results.subtitle = FALSE,
  centrality.plotting = FALSE,
  bf.message = FALSE,
  p.adjust.method = "holm",
  ggtheme = ggpubr::theme_pubr()
) +
  scale_color_manual(values = c(
    "dodgerblue2",
    "#E31A1C",
    "green4",
    "#6A3D9A"
  )) +
  labs(title = "Viral Load - Only SYMPTOMATIC Samples") +
  scale_y_log10())
```

```{r include = FALSE}
ggsave('fig2_viral_load_symptomatic.png',
       ct_N_stats_plot_sym)
```

```{r}
sessionInfo()
```