---
title: "Weekly COVID19 Positivity Rate"
author: "Stevin Wilson"
date: "12/21/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: no
    df_print: paged
    highlight: kate
  bookdown::pdf_document2:
    toc: yes
    latex_engine: lualatex
    number_sections: yes
    highlight: kate
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, include = FALSE}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(lubridate)
library(readxl)
library(DBI)
library(ggpubr)
library(plotly)
library(ggstatsplot)
library(car)
library(rstatix)
library(zoo)
library(kableExtra)
set.seed(100)
```

# Outline

This script is used to plot weekly positivity rate for COVID19 based on the diagnostics results from Clemson University's testing program.
```{r include = FALSE}
db_build <- dbConnect(RSQLite::SQLite(), "deidentifiedDB_v2021.12.08.sqlite")
sample_collection_table <- dbReadTable(db_build, "sample_collection")
dbDisconnect(db_build)
```
Samples are classified into different `order_priority` categories based on self-reported symptoms/exposure

```{r}
unique(sample_collection_table$order_priority)
```

```{r}
unique(sample_collection_table$performing_facility)
```

All possible categories for results status as as follows:

```{r}
unique(sample_collection_table$rymedi_result)
```

## Exclude Sample without `patient_id` or `collection_date`


```{r}
sample_collection_without_missing <- sample_collection_table %>%
  filter(
    !(is.na(collection_date) | is.na(patient_id)),
    performing_facility %in% c(
      "CLEMSON UNIVERSITY",
      "CLEMSON UNIVERSITY PERSONNEL",
      "CLEMSON COMMUNITY"
    ),
    rymedi_result %in% c("POSITIVE", "NEGATIVE")
  ) %>%
  mutate(
    collection_date = date(as_datetime(collection_date)),
    result_date = as_date(result_date),
    collection_week = lubridate::week(collection_date)
  ) %>%
  filter(collection_date >= "2021-01-01") %>%
  select(
    testkit_id, rymedi_result, collection_week, collection_date, population, order_priority, gender, patient_id
  ) %>%
  arrange(collection_date, rymedi_result)

glimpse(sample_collection_without_missing)
```

## Only retain the first COVID19 +ve sample from a patient in a period of 30 days

```{r}
positive_samples <- sample_collection_without_missing %>%
  filter(rymedi_result == "POSITIVE")

glimpse(positive_samples)
```

Number of such multi-test possibly-single-infection samples :
```{r}
multi_positive_patients <- positive_samples %>%
  group_by(patient_id) %>%
  filter(n() > 1) %>%
  arrange(patient_id, collection_date) %>%
  group_split()

length(multi_positive_patients)
```

```{r}
multi_positive_patients[1:2]
```

```{r}
testkit_ids_to_exclude <- c()

for (n in seq_along(multi_positive_patients)) {
  temp_tbl <- multi_positive_patients[[n]] %>%
    arrange(collection_date)
  diff_days <- diff(temp_tbl[["collection_date"]])

  for (m in seq_along(diff_days)) {
    if (diff_days[m] <= 30) {
      temp_c_date <- temp_tbl$testkit_id[m + 1]
      testkit_ids_to_exclude <- c(
        testkit_ids_to_exclude,
        temp_c_date
      )
    }
  }
}

glimpse(testkit_ids_to_exclude)
```

`sample_collection` table after excluding above `testkit_id`s

```{r}
sample_collection_without_repeats <- sample_collection_without_missing %>%
  filter(!(testkit_id %in% testkit_ids_to_exclude))

glimpse(sample_collection_without_repeats)
```

Patients with reinfection event with gap of more than 30 days between positive tests are retained :

```{r}
sample_collection_without_repeats %>%
  filter(rymedi_result == "POSITIVE") %>%
  group_by(patient_id) %>%
  filter(n() > 1) %>%
  arrange(patient_id, collection_date) %>%
  relocate(patient_id, collection_date) %>%
  select(-c(testkit_id))
```

Information regarding the week number for samples collection date

```{r}
week_dates_info <- sample_collection_without_missing %>%
  select(collection_week, collection_date) %>%
  distinct() %>%
  group_by(collection_week) %>%
  mutate(
    week_start = min(collection_date),
    week_end = max(collection_date),
    collection_week = as_factor(collection_week)
  ) %>%
  select(-c(collection_date)) %>%
  distinct()

glimpse(week_dates_info)
```

Positivity rate in different `order_priority` categories over the year 2021

```{r}
sample_collection_without_repeats %>%
  group_by(
    collection_week,
    rymedi_result,
    order_priority
  ) %>%
  summarize(count = n()) %>%
  mutate(
    count = replace(count, count == 0, 1),
    order_priority = factor(order_priority,
      levels = c(
        "SYMPTOMATIC",
        "ONE DAY",
        "EXPOSED",
        "SURVEILLANCE"
      )
    )
  ) %>%
  pivot_wider(
    names_from = rymedi_result,
    values_from = count
  ) %>%
  mutate(
    POSITIVE = replace(POSITIVE, is.na(POSITIVE), 0),
    NEGATIVE = replace(NEGATIVE, is.na(NEGATIVE), 0),
    TOTAL = NEGATIVE + POSITIVE,
    POSITIVITY = round((POSITIVE / TOTAL) * 100, 2),
    collection_week = as_factor(collection_week)
  ) %>%
  full_join(week_dates_info,
    by = "collection_week"
  ) %>%
  relocate(
    collection_week, week_start, week_end, order_priority,
    TOTAL, POSITIVE, NEGATIVE, POSITIVITY
  ) %>%
  arrange(collection_week, order_priority) %>%
  kbl() %>%
  kable_classic_2(
    full_width = F,
    latex_options = c(
      "hold_position",
      "striped"
    )
  )
```

```{r}
fig_data <- sample_collection_without_repeats %>%
  filter(order_priority %in% c(
    "SYMPTOMATIC",
    "EXPOSED",
    "SURVEILLANCE"
  )) %>%
  group_by(
    collection_week,
    rymedi_result,
    order_priority
  ) %>%
  summarize(count = n()) %>%
  mutate(
    count = replace(count, count == 0, 1),
    order_priority = factor(order_priority,
      levels = c(
        "SYMPTOMATIC",
        "EXPOSED",
        "SURVEILLANCE"
      )
    )
  ) %>%
  pivot_wider(
    names_from = rymedi_result,
    values_from = count
  ) %>%
  mutate(
    POSITIVE = replace(POSITIVE, is.na(POSITIVE), 0),
    NEGATIVE = replace(NEGATIVE, is.na(NEGATIVE), 0),
    TOTAL = NEGATIVE + POSITIVE,
    POSITIVE = replace(POSITIVE, POSITIVE == 0, 1),
    TOTAL = replace(TOTAL, TOTAL == 0, 1)
  ) %>%
  select(-c(NEGATIVE)) %>%
  pivot_longer(!c(
    collection_week,
    order_priority
  ),
  names_to = "STATUS", values_to = "COUNT"
  ) %>%
  mutate(STATUS = factor(STATUS,
    levels = c(
      "TOTAL",
      "POSITIVE"
    )
  ))

glimpse(fig_data)
```

```{r message = FALSE, warning = FALSE, fig.width= 7, fig.height= 5}
p <- fig_data %>%
  mutate(
    STATUS = as.character(STATUS),
    STATUS = dplyr::recode(STATUS,
      `POSITIVE` = "Number of Positive Cases",
      `TOTAL` = "Total Number of Tests"
    ),
    STATUS = fct_rev(STATUS)
  ) %>%
  ggbarplot(
    x = "collection_week",
    y = "COUNT",
    fill = "STATUS",
    alpha = 0.8,
    sorting = "none",
    format.scale = TRUE,
    yscale = "log10",
    title = "New COVID-19 Cases in the Clemson Area",
    subtitle = "2021 Data",
    xlab = "Week Number",
    ylab = "Count",
    position = position_dodge(preserve = "single")
  )

p <- facet(p,
  facet.by = c("order_priority"),
  nrow = 3
)

ggpar(p,
  legend = "bottom",
  legend.title = "",
  ggtheme = ggstatsplot::theme_ggstatsplot()
)
```

```{r}
fig_data <- fig_data %>%
  left_join(week_dates_info %>%
    mutate(collection_week = as.double(collection_week)),
  by = "collection_week"
  ) %>%
  relocate(collection_week, week_start, week_end) %>%
  mutate(
    STATUS = as.character(STATUS),
    STATUS = dplyr::recode(STATUS,
      `POSITIVE` = "Number of Positive Cases",
      `TOTAL` = "Total Number of Tests"
    ),
    STATUS = fct_rev(STATUS)
  )

fig_data %>%
  arrange(collection_week, order_priority, STATUS) %>%
  kbl() %>%
  kable_classic_2(
    full_width = F,
    latex_options = c(
      "hold_position",
      "striped"
    )
  )
```

```{r}
fig_data %>%
  write_csv("data_for_fig3.csv")
```
